---
title: 'Fertility and Wealth: Causal Evidence for Latin America'
author: "Edison Choque Sanchez"
date: "30/06/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_download: yes
    theme: flatly
  pdf_document:
    toc: yes
subtitle: Work in progress
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

In this file we show the process to estimate the causal effect from fertility to household wealth. This process has three stages. First, we import the data base and explore it, but also we generate a subset of the data base because women are the group of analysis. Second, all the variable of interest should be created. For instance most of them should be transformed into the correct format. Finally, we show the result according to the model.

# 2. DHS program data

DHS surveys collect primary data using several types of questionnaires. A household questionnaire is used to collect information on characteristics of the household's dwelling unit, and data related to the height and weight for women and children in the household. It is also used to identify members of the household who are eligible for an individual interview. Eligible respondents are then interviewed using an individual questionnaire.

In a majority of DHS surveys eligible individuals include women of reproductive age (15-49) and men age 15-59, or in some cases 15-54. In some countries only women are interviewed. Individual questionnaires include information on fertility, family planning and maternal and child health.

We use the individual record data. It mean that all women between 15 - 49 years old were interviewed. You can visit the following website to know about DHS program and data. 
[DHS program](https://dhsprogram.com/data/Getting-Started.cfm){target="_blank"}

```{r include=FALSE}
library(readstata13)
library(survey)
library(srvyr)
library(dplyr)
library(tidyverse)
library(sem)
library(cragg) # for Stock - Yogo test for weak instruments
```

## Import data

```{r include=FALSE}
setwd("J:/Mi unidad/Consultorias/Gus_Canavire/ProyectoDHS/Data/phase 5")
dir()
base_model1 <- read.dta13("base_ir1.DTA")
base_model2 <- read.dta13("base_ir2.DTA")
```

In order to study specific data. First, we use two databases from 6 Latin American countries. These are Bolivia, Colombia, Republica Dominicana, Guyana, Honduras, and Haiti. We collect the data between 2005 to 2009 according to the phase 5 - DHS program. 

We develop two econometric models. Both are causal models to measure raising children's impact on household wealth.

To know the definition for the most of the variables, we suggest downloading the pdf files from the following link: [Files](https://dhsprogram.com/publications/publication-dhsg4-dhs-questionnaires-and-manuals.cfm){target="_blank"}. Once you enter to the link, download two pdf files: DHS-V Recode Manual and DHS-V Recode Map. In both,  you can find more information about the variables and data structure. 

The first econometric model considers the "infertility shock" as the instrument. The second one uses the "twins" variable as an alternative instrument. Most of them are changed to the correct format. We use both the **base_model1** and **base_model2** for the first and second econometric models respectively

This presentation is divided into two parts. First, we show the description of the variables. Second, we explain the principal results of the econometrics models.

# 1. Description of the Database

## Model 1: Infertility shock

First, we explore the dependent variable which is the Wealth index factor score
 
```{r descriptive 1.1}
hist(base_model1$score_wi)
summary(base_model1$score_wi)
```

Second, we describe the endogenous variable that is fertility, which is the number of living children. But also, we can define this variable in different ways. See some of these options:

```{r descriptive 1.2}
table(base_model1$children_home) # Number of children living at home 
table(base_model1$living_children) # Number of living children
table(base_model1$v137) #v137 -Number of children 5 and under
table(base_model1$v219) # v219 -Living children + current preg
```

Third, we describe the instrument for the endogenous variable. This instrument is defined in two ways. Both are self-declarations of infertility for women.

a) when the question is: Main reason not to use meth, and the answer is Subfecund, infecund. 

b) when the question is: Fertility preference, and the answer is: Declared infecund. 

According to these results, on average the number of living children in the infertile group is lower than the fertile group, so that makes sense.

```{r descriptive 1.3}
base_model1 %>% group_by(infertility1) %>%  summarise(mean = mean(living_children), n = n())
base_model1 %>% group_by(infertility2) %>%  summarise(mean = mean(living_children), n = n())
base_model1 %>% group_by(instrument) %>%  summarise(mean = mean(living_children), n = n()) #instrument is defined as infertility1 or  infertility2
```

Finally, we describe the rest of the variables for model 1.

```{r descriptive 1.4}
summary(base_model1$v012) #V012 - Current age - respondent
summary(base_model1$v212) # v212 Age of respondent at 1st birth
summary(base_model1$v133) # v133 - Education in single years
table(base_model1$urban) # v025 - Type of place of residence
```

### Model 2: Twins 

In this econometric model, the outcome variable is the same as in model 1. Also, the endogenous variable is the same. But in this case, the database changed because we consider households with equal or more than one child. 

Now we can describe the instrumental variable for model 2. This is defined as *twin*. We assume that mothers with the experience of having twins have to raise extra children. That is why it works as a good instrument to measure the causal effect. 

In the next chunk, we show a summary table for the mean of the number of children living at home according to mothers who had twins and the ones that did not. The result shows that this mean is lower for the mothers who never had twins than the ones who did, so that makes sense.

```{r descriptive 2.1}
base_model2 %>% group_by(twin) %>%  summarise(mean = mean(children_home), n = n())
```

Finally, we present the description for the rest of the variables for model 2.

```{r}
summary(base_model2$max_agechildren) # Maximun age of the children at home
summary(base_model2$v012) #V012 - Current age - respondent
summary(base_model2$v212) # v212 Age of respondent at 1st birth
summary(base_model2$v133) # v133 - Education in single years
table(base_model2$urban) # 1 urbano - 0 rural
```

# 4. Econometrics Models

## Model 1: Infertility shock

Definition of the dependent variables:

 - score_wi: Wealth index factor score standardized
 
Definition of the endogenous variables:
 
 - living_children: number of living children
 
The covariates were defined as follow:

 - V012: Current age - respondent
 - v212: Age of respondent at 1st birth
 - v133: Education in single years
 - urban: the responded live in urban area
 
```{r}
model_a1 <- lm(score_wi ~ living_children + v012 + v133 + v212 + urban, data = base_model1)
summary(model_a1)
confint(model_a1)
```

For the model_a1, we do not consider any instrument, so it will be useful to compare these results with the model_a2 and model_a3 that apply the infertility instrument. 

```{r}
model_a2 <- tsls(score_wi~ living_children + v012 + v133 + v212 + urban, 
                 ~ instrument + v012 + v133 + v212 + urban, data = base_model1)
summary(model_a2)
confint(model_a2)
```

```{r}
model_a3 <- tsls(score_wi~ living_children + v012 + v133 + v212, 
                 ~ instrument + v012 + v133 + v212, data = base_model1)
summary(model_a3)
confint(model_a3)
```

The principal results show that the causal effect from fertility to the score of wealth index in negative for the model_a3, and it is statistically significant. It is important to notice that the causal effect from the *living_children* variable is small in average for each children. And also, notice that is we control by *urban* variable, the causal effect is not statistically significant (see model_2a). 


## Model 2: Twins

Definition of the dependent variables:

 - score_wi: Wealth index factor score standardized
 
Definition of the endogenous variables:
 
 - children_home: number of children living at home
 
The covariates were defined as follow:

 - V012: Current age - respondent
 - v212: Age of respondent at 1st birth
 - v133: Education in single years
 - urban: the responded live in urban area
 
```{r}
model_b1 <- lm(score_wi ~ children_home + v012 + v133 + v212 + urban, 
               data = base_model2)
summary(model_b1)
confint(model_b1)
```

For the model_b1, we do not consider any instrument, so it will be useful to compare these results with the model_b2 and model_b3 that apply the twin instrument. 

```{r}
model_b2 <- tsls(score_wi~ children_home + v012 + v133 + v212  + urban, 
                 ~ twin + v012 + v133 + v212  + urban , data = base_model2)
summary(model_b2)
confint(model_b2)
```

The principal result shows that the casual effect from fertility over the score of wealth index is statistically significant for the model_b1. But if we delete the variable *urban*, the causal effect is statistically significant and bigger. see the result:

```{r}
model_b3 <- tsls(score_wi~ children_home + v012 + v212, 
                 ~ twin + v012 + v212 , data = base_model2)
summary(model_b3)
confint(model_b3)
```


# 5. Discussion and recomendations

The causal effect from fertility over the score of wealth index is negative for the model_a2, model_a3, model_b2 and model_b3. So, we can conclude  that the causal effect is statistically significant  for each child. 

Also, the weak 

There are three aspect to consider in order to improve this preliminary results: 

 - Check if the results change if we use individual survey weight
 
 - We should consider the tests for weak instrument (we develop this test in Stata)
 
 - We should consider the test for the external validation like Hansen test
 
 - We can use a restriction for the maximum age of the children living at home. 

