---
title: 'Fertility and Wealth: Causal Evidence'
author: "Edison Choque Sanchez"
date: "30/5/2022"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_download: yes
    theme: flatly
  pdf_document:
    toc: yes
subtitle: Work in progress
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

In this file we show the process to estimate the causal effect from fertility to household wealth. This process has three stages. First, we import the data base and explore it, but also we generate a subset of the data base because women are the group of analysis. Second, all the variable of interest should be created. For instance most of them should be transformed into the correct format. Finally, we show the result according to the model.

# 2. Data Generation

DHS surveys collect primary data using several types of questionnaires. A household questionnaire is used to collect information on characteristics of the household's dwelling unit, and data related to the height and weight for women and children in the household. It is also used to identify members of the household who are eligible for an individual interview. Eligible respondents are then interviewed using an individual questionnaire.

In a majority of DHS surveys eligible individuals include women of reproductive age (15-49) and men age 15-59, or in some cases 15-54. In some countries only women are interviewed. Individual questionnaires include information on fertility, family planning and maternal and child health.

We use the individual record data. It mean that all women between 15 - 49 years old were interviewed. You can visit the following website to know about DHS program and data. 
[DHS program](https://dhsprogram.com/data/Getting-Started.cfm){target="_blank"}

## Libraries used are:

```{r include=FALSE}
library(readstata13)
library(survey)
library(srvyr)
library(dplyr)
library(tidyverse)
library(sem)
```

## Import data

```{r}
setwd("J:/Mi unidad/Consultorias/Gus_Canavire/ProyectoDHS/Data/2008")
dir()
bdd_ir <- read.dta13("BOIR51FL.DTA", convert.factors = T)
```

## Data generation

In order to study and specific data, First, we select all variables that we consider to study.  Second, we choose all the observation related with the female head of the household or the wife. 

The selection of the principal variables is described in the next chunk. But also, is described how we select only the observations for wife or female head. 

```{r}
ir1 <-  select(bdd_ir, caseid,v002,v003, v005, v012, v013, v133, v135, v137,
               v139, v150, v151,
               v152,v201, v219, v376, v605, v602,v616,
               v191,v190, v218, v731)
ir1 <- filter(ir1, v150=="wife or husband" | v150=="head")
```

To know the definition of these variables, we suggest to download the pdf files from the following link: [Files](https://dhsprogram.com/publications/publication-dhsg4-dhs-questionnaires-and-manuals.cfm){target="_blank"}. Once you enter to the link, download two pdf files: DHS-V Recode Manual and DHS-V Recode Map. In both you can find more information about the variables and data structure. 

# 3. Creation of variables 

the following variables are created for the 2SLS model. Most of them are changed to the correct format. 

```{r}
ir1<- ir1 %>% mutate(infertility1 = v376 %in% 24) #v376 - Main reason not to use a meth.
ir1$infertility1_a=as.numeric(ir1$infertility1)

ir1<- ir1 %>% mutate(infertility2 = v602 %in% 5) # v602 - Fertility preference
ir1$infertility2_a=as.numeric(ir1$infertility2)

ir1$children <-  recode(ir1$v218, `0`=0L, `1`=1L, `2`=1L, `3`=1L, `4`=1L, `5`=1L, 
                        `6`=1L, `7`=1L, `8`=1L,`9`=1L, `10`=1L,`11`=1L,
                        `12`=1L, `13`=1L, `14`=1L) # v218- Number of living children

v2 <- c("poorest", "poorer")
ir1<- ir1 %>% mutate(wealth = v190 %in% v2) # v190 - Wealth index
ir1$wealth_a<- as.numeric(ir1$wealth) # wealth_a -  1 "poorest - poorer" 0 is "middle - richer- richest"

m1 <- c("la paz", "oruro", "potosí")
m2 <- c("cochabamba", "chuquisaca", "tarija")
m3 <- c("pando", "beni", "santa cruz")
ir1 <- ir1 %>% mutate(altiplano =v139 %in% m1, 
      valle =v139 %in% m2, llano=v139 %in% m3) # v139 - De jure region of residence
ir1$altiplano_a<- as.numeric(ir1$altiplano)
ir1$valle_a<- as.numeric(ir1$valle)
ir1$llano_a<- as.numeric(ir1$llano)
```

# 4. Description of the **ir1** data base

First, we explore the dependent variable that is the Wealth index factor score, but also the binary variable *wealth_a*. 
 
```{r descriptive 1}
ir1 <- ir1 %>% rename(score_wi=v191) #v191 - Wealth index factor score
hist(ir1$score_wi)
summary(ir1$score_wi)
table(ir1$wealth_a) 
```

Second, we describe the endogenous variable that is fertility, that is number of living children. We can define this variable in different ways. See some of these options:

```{r descriptive 2}
table(ir1$children)
table(ir1$v218) 
table(ir1$v137) #v137 -Number of children 5 and under
table(ir1$v219) # v219 -Living children + current preg
```

Third, we describe the instrument for the endogenous variable. This instrument is defined in two ways. Both are self declaration of infertility. 

a) when the question is: Main reason not to use a meth, and the answer is: Subfecund, infecund. 

b) when the question is: Fertility preference, and the answer is: Declared infecund. 

According to these results, in average the number of children in the infertile group is lower than the fertile group, that make sense. 

```{r descriptive 3}
ir1 %>% group_by(v013,infertility1_a) %>%  summarise(mean = mean(v137), n = n()) # v013 - Age 5-year groups
ir1 %>% group_by(v013,infertility2_a) %>%  summarise(mean = mean(v137), n = n())
```

Finally, we describe the rest of the variable for the model. 

```{r descriptive 4}
summary(ir1$v012) #V012 - Current age - respondent
summary(ir1$v133) # v133 - Education in single years
table(ir1$altiplano_a)
table(ir1$valle_a)
table(ir1$llano_a)
```

# 5. Econometric Model

Definition of the dependent variables:

 - score_wi : Wealth index factor score
 
 - wealth_a: 1 is "poorest - poorer" and 0 is "middle - richer- richest"
 
Definition of the endogenous variables:
 
 - children: 1 with children and 0 without children
 
 - v219: Living children + current preg
 
 - v137: Number of children 5 and under
 
The covariates were defined above. 
 
```{r}
model0 <- lm(score_wi ~ v219 + v012 + v133 + altiplano_a + valle_a + llano_a, 
             data = ir1)
summary(model0)
confint(model0)

model1 <- tsls(score_wi~ children + v012 + v133 + altiplano_a + valle_a + llano_a, ~ infertility1_a + v012 + v133 + altiplano_a + valle_a + llano_a, data = ir1)
summary(model1)
confint(model1) 

model2 <- tsls(score_wi~ v219+ v012 + v133 + altiplano_a + valle_a + llano_a, 
               ~ infertility1_a+ v012 + v133 + altiplano_a + valle_a + llano_a , data = ir1)
summary(model2)
confint(model2)

model3 <- tsls(score_wi~ v137+ v012 + v133 + altiplano_a + valle_a + llano_a,
               ~ infertility1_a + v012 + v133 + altiplano_a + valle_a + llano_a  , data = ir1)
summary(model3)
confint(model3)
  
model4 <- tsls(wealth_a~ children + v012 + v133 + altiplano_a + valle_a + llano_a, ~ infertility1_a + v012 + v133 + altiplano_a + valle_a + llano_a, data = ir1)
summary(model4)
confint(model4) 

model5 <- tsls(wealth_a~ v219+ v012 + v133 + altiplano_a + valle_a + llano_a, 
               ~ infertility1_a+ v012 + v133 + altiplano_a + valle_a + llano_a , data = ir1)
summary(model5)
confint(model5)

model6 <- tsls(wealth_a~ v137+ v012 + v133 + altiplano_a + valle_a + llano_a,
               ~ infertility1_a + v012 + v133 + altiplano_a + valle_a + llano_a  , data = ir1)
summary(model6)
confint(model6)
```

# 6. Discussion and recomendations

The causal effect from fertility is negative over the score of the wealth index. On the other hand, there is no evidence when we consider the wealth index as binary variable. Finally, if we just run a lineal regression, the effect is negative over the wealth index in the case of the variable: v219 - Living children + current preg. But this effect higher (see model0 above).

There are three aspect to consider in order to improve this work. 

 - Check if keep the Wealth index factor score is the best option
 
 - We should consider the test for weak instrument 
 
 - We should consider the test for the external validation like Hansen test
 
 - We can consider the second instrument: infertility2_a for all the models. But it does not seems to have a big change because of the definition is similar in comparison to infertility1_a
 
 - We can prove the twins instrument

